---
title: Projectfinal
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warnings = FALSE, 
  messages = FALSE, 
  prompt = FALSE,
  fig.path = "man/figures/README-",
  out.width = "100%"
)

```



<!-- badges: start -->

<!-- badges: end -->

This R package is designed to provide a template for users to conduct various statistical tests, including the two sample t-test, simple linear regression test and chi-squared test for independence, on any sample data set to solve the research questions of interest.

## First things first

`projectfinal` is a toy project and is still under development. You can install the latest version from [GitHub](https://github.com/) with:

```{r, install, results = "hide", eval = FALSE}
# install.packages("remotes")
remotes::install_github("hemishahuja140/mq-stat1378-project2022")
```

## Take it for a spin

`projectfinal` contains three wrapper functions with the three seperate tests. For our example, we will consider `mytest_chi` which will perform a chi-squared test for independence.

```{r example}
devtools::install_github("https://github.com/hemishahuja140/practice")

library("projectfinal")
 
mytest_chi(project2022)
```


